---
title: "coding challenge 7"
author: "lihui xiang"
date: "2025-04-03"
output: pdf_document
---
####11. Read in the data called “PlantEmergence.csv” using a relative file path and load the following libraries. tidyverse, lme4, emmeans, multcomp, and multcompView. Turn the Treatment , DaysAfterPlanting and Rep into factors using the function as.factor
```{r}
library(tidyverse)
library(lme4)
library(emmeans)
library(multcomp)
library(multcompView)
STAND <- read.csv("PlantEmergence.csv")   
##Turn the Treatment , DaysAfterPlanting  and Rep into factors using the function as.factor
STAND$Treatment <- as.factor(STAND$Treatment)
STAND$DaysAfterPlanting <- as.factor(STAND$DaysAfterPlanting)
STAND$Rep <- as.factor(STAND$Rep)
```
###2. Fit a linear model to predict Emergence using Treatment and DaysAfterPlanting along with the interaction. Provide the summary of the linear model and ANOVA results
```{r}
lm_model <- lm(Emergence ~ Treatment * DaysAfterPlanting, data = STAND)
summary(lm_model)
anova(lm_model)
```
###3.Based on the results of the linear model in question 2, do you need to fit the interaction term? Provide a simplified linear model without the interaction term but still testing both main effects. Provide the summary and ANOVA results. Then, interpret the intercept and the coefficient for Treatment 2.
```{r}
###we do not need to fit the interaction term as the interaction term is not significant.
# Fit a simplified linear model without the interaction term
lm_simple <- lm(Emergence ~ Treatment + DaysAfterPlanting, data = STAND)

# Display the summary of the simplified linear model
summary(lm_simple)

# Perform ANOVA to test the significance of main effects
anova(lm_simple)

#The intercept (182.25) represents the expected emergence when the treatment is at level 1 (reference level) and DaysAfterPlanting is at its reference level. The coefficient for Treatment 2 (-136.50) means that Treatment 2 has 136.5 fewer emergences compared to Treatment 1.
```
###4. Calculate the least square means for Treatment using the emmeans package and perform a Tukey separation with the compact letter display using the cld function. Interpret the results.
```{r}
lsmeans_treat <- emmeans(lm_simple, ~ Treatment)
results_cld <- cld(lsmeans_treat, alpha = 0.05, Letters = letters)
print(results_cld)
```
##5.The provided function lets you dynamically add a linear model plus one factor from that model and plots a bar chart with letters denoting treatment differences. Use this model to generate the plot shown below. Explain the significance of the letters. 
```{r}
plot_cldbars_onefactor <- function(lm_model, factor) {
  data <- lm_model$model
  variables <- colnames(lm_model$model)
  dependent_var <- variables[1]
  independent_var <- variables[2:length(variables)]

  lsmeans <- emmeans(lm_model, as.formula(paste("~", factor))) # estimate lsmeans 
  Results_lsmeans <- cld(lsmeans, alpha = 0.05, reversed = TRUE, details = TRUE, Letters = letters) # contrast with Tukey adjustment by default.
  
  # Extracting the letters for the bars
  sig.diff.letters <- data.frame(Results_lsmeans$emmeans[,1], 
                                 str_trim(Results_lsmeans$emmeans[,7]))
  colnames(sig.diff.letters) <- c(factor, "Letters")
  
  # for plotting with letters from significance test
  ave_stand2 <- lm_model$model %>%
    group_by(!!sym(factor)) %>%
    dplyr::summarize(
      ave.emerge = mean(.data[[dependent_var]], na.rm = TRUE),
      se = sd(.data[[dependent_var]]) / sqrt(n())
    ) %>%
    left_join(sig.diff.letters, by = factor) %>%
    mutate(letter_position = ave.emerge + 10 * se)
  
  plot <- ggplot(data, aes(x = !! sym(factor), y = !! sym(dependent_var))) + 
    stat_summary(fun = mean, geom = "bar") +
    stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.5) +
    ylab("Number of emerged plants") + 
    geom_jitter(width = 0.02, alpha = 0.5) +
    geom_text(data = ave_stand2, aes(label = Letters, y = letter_position), size = 5) +
    xlab(as.character(factor)) +
    theme_classic()
  
  return(plot)
}
plot1 <- plot_cldbars_onefactor(lm_simple, "Treatment")
print(plot1)
##This plot shows the average number of emerged plants for each treatment, with error bars showing how much the values vary. The letters on top come from Tukey’s test — if two treatments share the same letter, it means there’s no real difference between them. If they have different letters, they’re significantly different. For example, Treatment 2 had way fewer emerged plants than the others, while Treatments 1, 3, 4, and 5 were pretty similar.
```

###6.You can view the GitHub folder for this challenge here:  
[GitHub - coding_challenge7](https://github.com/Siya2904-cell/Lihui-reproducibility/tree/main/coding_challenge7)